### 2018-07-29 other_R_code/article_selection_process.Rnw
write up results

\documentclass{article}

\begin{document}

We only have access to a big data set via boolean search.

Sub-sets of the data set returned by a single key word on a single day are large.For example for the single search term `election' for the single day of 1 August 1868 returns 1982 articles. The search term `riot' for the same day (close to the 1868 General Election) returns 360 articles, 228 of which do not also contain the word election.

With even a small number of search terms it quicly becomes impractical to examine all the documents even for a single day.

Following King et al we define
$S$ - the search set of all documents in the British Newspaper Archive
$T$ - the target set of all documents in the British Newspaper Archive which are about election violence
$R$ - a reference set of documents which are about election violence

The task is to identify $T$ from $S$ in a form where $T$ can be

It is trivial to define an algorithm which obtains a subset of $S$ which contains $T$, because $S \subseteq S$ and $T \subset S$. Algorithms which aim to maximise the chances of obtaining all of $T$ will tend to return $S$.

Our task is to find a good method for returning $T$ from $S$ in a form which we can analyse. By a \textit{good} method we mean a method which returns a greater proportion of $T$, and a greater ratio of $T$ to $\neg T$ than alternative methods. The main alternative method is manual searching by historians.

\section{Keyword Identification}
How difficult is it to find search terms which return all of

\end{document}