### 2018-07-30 other_R_code/article_selection_process.Rnw
further result classification write-up

\documentclass{article}

\begin{document}
<<loadpackages, echo=FALSE, warning=FALSE, message=FALSE>>=
library(durhamevp)
library(tidyverse)
@

We only have access to a big data set via boolean search.

Sub-sets of the data set returned by a single key word on a single day are large.For example for the single search term `election' for the single day of 1 August 1868 returns 1982 articles. The search term `riot' for the same day (close to the 1868 General Election) returns 360 articles, 228 of which do not also contain the word election.

With even a small number of search terms it quicly becomes impractical to examine all the documents even for a single day.

Following King et al we define
$S$ - the search set of all documents in the British Newspaper Archive
$T$ - the target set of all documents in the British Newspaper Archive which are about election violence
$R$ - a reference set of documents which are about election violence

The task is to identify $T$ from $S$ in a form where $T$ can be

It is trivial to define an algorithm which obtains a subset of $S$ which contains $T$, because $S \subseteq S$ and $T \subset S$. Algorithms which aim to maximise the chances of obtaining all of $T$ will tend to return $S$.

Our task is to find a good method for returning $T$ from $S$ in a form which we can analyse. By a \textit{good} method we mean a method which returns a greater proportion of $T$, and a greater ratio of $T$ to $\neg T$ than alternative methods. The main alternative method is manual searching by historians.

\section{The data}
<<loaddata, cache=TRUE, tidy=TRUE>>=
classdocs<-durhamevp::get_classified_docs()
classified_corpus<-quanteda::corpus(classdocs[,c("fakeid", "ocr", "EV_article")], text_field = "ocr")
classified_dfm <- preprocess_corpus(classified_corpus, stem=FALSE, min_termfreq=20, min_docfreq = 10)
@

In the data there are \Sexpr{nrow(classdocs)} cases:
\begin{itemize}
  \item \textbf{\Sexpr{sum(classdocs$election_article==0)}} non-election articles.
  \item \textbf{\Sexpr{sum(classdocs$EV_article==1)}} election violence articles.
  \item \textbf{\Sexpr{sum(classdocs$just_election==1)}} election (but not violent) articles.
\end{itemize}

\section{Keyword Identification}
Note it is important to make keyword identification somewhat selective of $T$ from $S$ otherwise even very good stage 2 \& 3 selection processes the false positives will overwhelm the true positives.

Algorithm:
\begin{enumerate}
  \item use classifier on $R$ and $S$ to identify two lists of keywords
  \item generate probability from classifier parameters
  \item add to keyword list based on probability
\end{enumerate}
<<findkeywords>>=
keywords<-nb_keywords(classified_dfm, "EV_article")
knitr::kable(head(keywords, 20))
@

\section{Refinement Using Description}

<<usingdescription>>=
description_corpus<-quanteda::corpus(classdocs[,c("fakeid", "description", "EV_article")], text_field = "description")
description_dfm <- preprocess_corpus(description_corpus, stem=FALSE, min_termfreq=5, min_docfreq = 2)

both_dfms<-split_dfm(description_dfm, n_train = 1500)
nb<-quanteda::textmodel_nb(both_dfms$train, y=quanteda::docvars(both_dfms$train, "EV_article"), prior="uniform")
prob_nb<-predict(nb, newdata = both_dfms$test, type="probability")
pred_nb<-data.frame(predict(nb, newdata = both_dfms$test, type="class"))
res<-data.frame(predict_nb=pred_nb, prob_nb)
names(res)<-c("predict_nb", "prob_notev", "prob_ev")
assess_classification(organise_results(both_dfms$test, res)) %>%
  filter(rowname %in% c("Precision", "Recall", "F1")) %>%
  kable()
@

\section{Refinement Using OCR}

<<usingocr>>=
ocr_corpus<-quanteda::corpus(classdocs[,c("fakeid", "ocr", "EV_article")], text_field = "ocr")
ocr_dfm <- preprocess_corpus(ocr_corpus, stem=FALSE, min_termfreq=5, min_docfreq = 2)

both_dfms<-split_dfm(ocr_dfm, n_train = 1500)
nb<-quanteda::textmodel_nb(both_dfms$train, y=quanteda::docvars(both_dfms$train, "EV_article"), prior="uniform")
prob_nb<-predict(nb, newdata = both_dfms$test, type="probability")
pred_nb<-data.frame(predict(nb, newdata = both_dfms$test, type="class"))
res<-data.frame(predict_nb=pred_nb, prob_nb)
names(res)<-c("predict_nb", "prob_notev", "prob_ev")
assess_classification(organise_results(both_dfms$test, res)) %>%
  filter(rowname %in% c("Precision", "Recall", "F1")) %>%
  kable()
@

\end{document}