### 2018-07-31 other_R_code/article_selection_process.tex

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

We only have access to a big data set via boolean search.

Sub-sets of the data set returned by a single key word on a single day are large.For example for the single search term `election' for the single day of 1 August 1868 returns 1982 articles. The search term `riot' for the same day (close to the 1868 General Election) returns 360 articles, 228 of which do not also contain the word election.

With even a small number of search terms it quickly becomes impractical to examine all the documents even for a single day.

Following King et al we define
$S$ - the search set of all documents in the British Newspaper Archive
$T$ - the target set of all documents in the British Newspaper Archive which are about election violence
$R$ - a reference set of documents which are about election violence

The task is to identify $T$ from $S$ in a form where $T$ can be

It is trivial to define an algorithm which obtains a subset of $S$ which contains $T$, because $S \subseteq S$ and $T \subset S$. Algorithms which aim to maximise the chances of obtaining all of $T$ will tend to return $S$.

Our task is to find a good method for returning $T$ from $S$ in a form which we can analyse. By a \textit{good} method we mean a method which returns a greater proportion of $T$, and a greater ratio of $T$ to $\neg T$ than alternative methods. The main alternative method is manual searching by historians.

\section{The data}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{classdocs} \hlkwb{<-} \hlstd{durhamevp}\hlopt{::}\hlkwd{get_classified_docs}\hlstd{()}
\hlcom{# just some candidate document for speed/space reasons}
\hlstd{unclassdocs} \hlkwb{<-} \hlstd{durhamevp}\hlopt{::}\hlkwd{get_candidate_documents}\hlstd{(}\hlnum{3000}\hlopt{:}\hlnum{6000}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

In the data there are 1894 cases:
\begin{itemize}
  \item \textbf{703} non-election articles.
  \item \textbf{769} election violence articles.
  \item \textbf{422} election (but not violent) articles.
\end{itemize}

\section{Keyword Identification}
Note it is important to make keyword identification somewhat selective of $T$ from $S$ otherwise even very good stage 2 \& 3 selection processes the false positives will overwhelm the true positives.

Algorithm:
\begin{enumerate}
  \item use classifier on $R$ and $S$ to identify two lists of keywords
  \item generate probability from classifier parameters
  \item add to keyword list based on probability
\end{enumerate}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{classified_corpus}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{corpus}\hlstd{(classdocs[,}\hlkwd{c}\hlstd{(}\hlstr{"fakeid"}\hlstd{,} \hlstr{"ocr"}\hlstd{,} \hlstr{"EV_article"}\hlstd{)],} \hlkwc{text_field} \hlstd{=} \hlstr{"ocr"}\hlstd{)}
\hlstd{classified_dfm} \hlkwb{<-} \hlkwd{preprocess_corpus}\hlstd{(classified_corpus,} \hlkwc{stem}\hlstd{=}\hlnum{FALSE}\hlstd{,} \hlkwc{min_termfreq}\hlstd{=}\hlnum{20}\hlstd{,} \hlkwc{min_docfreq} \hlstd{=} \hlnum{10}\hlstd{)}

\hlstd{keywords}\hlkwb{<-}\hlkwd{nb_keywords}\hlstd{(classified_dfm,} \hlstr{"EV_article"}\hlstd{)}
\hlstd{knitr}\hlopt{::}\hlkwd{kable}\hlstd{(}\hlkwd{head}\hlstd{(keywords,} \hlnum{20}\hlstd{))}
\end{alltt}
\end{kframe}
\begin{tabular}{l|l|r|r|r}
\hline
  & rowname & 0 & 1 & id\\
\hline
12299 & roughly & 0.0277149 & 0.9722851 & 12299\\
\hline
12300 & riotously & 0.0315492 & 0.9684508 & 12300\\
\hline
12280 & smashing & 0.0336163 & 0.9663837 & 12280\\
\hline
12298 & bicycle & 0.0347548 & 0.9652452 & 12298\\
\hline
12302 & pontypool & 0.0366149 & 0.9633851 & 12302\\
\hline
11621 & smashed & 0.0468791 & 0.9531209 & 11621\\
\hline
12303 & bedminster & 0.0499935 & 0.9500065 & 12303\\
\hline
12279 & p.c & 0.0528924 & 0.9471076 & 12279\\
\hline
12289 & roughs & 0.0548967 & 0.9451033 & 12289\\
\hline
12257 & rioters & 0.0579995 & 0.9420005 & 12257\\
\hline
12044 & stoned & 0.0585510 & 0.9414490 & 12044\\
\hline
12301 & abersychan & 0.0585510 & 0.9414490 & 12301\\
\hline
12304 & lofts & 0.0585510 & 0.9414490 & 12304\\
\hline
12295 & missile & 0.0611686 & 0.9388314 & 12295\\
\hline
11754 & supt & 0.0744892 & 0.9255108 & 11754\\
\hline
11959 & foley & 0.0744892 & 0.9255108 & 11959\\
\hline
11977 & missiles & 0.0776601 & 0.9223399 & 11977\\
\hline
9149 & staves & 0.0819225 & 0.9180775 & 9149\\
\hline
12094 & panes & 0.0835908 & 0.9164092 & 12094\\
\hline
12268 & terrell & 0.0835908 & 0.9164092 & 12268\\
\hline
\end{tabular}

\end{knitrout}

\section{Refinement Using Description}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{description_corpus}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{corpus}\hlstd{(classdocs[,}\hlkwd{c}\hlstd{(}\hlstr{"fakeid"}\hlstd{,} \hlstr{"description"}\hlstd{,} \hlstr{"EV_article"}\hlstd{)],} \hlkwc{text_field} \hlstd{=} \hlstr{"description"}\hlstd{)}
\hlstd{description_dfm} \hlkwb{<-} \hlkwd{preprocess_corpus}\hlstd{(description_corpus,} \hlkwc{stem}\hlstd{=}\hlnum{FALSE}\hlstd{,} \hlkwc{min_termfreq}\hlstd{=}\hlnum{5}\hlstd{,} \hlkwc{min_docfreq} \hlstd{=} \hlnum{2}\hlstd{)}

\hlstd{both_dfms}\hlkwb{<-}\hlkwd{split_dfm}\hlstd{(description_dfm,} \hlkwc{n_train} \hlstd{=} \hlnum{1000}\hlstd{)}

\hlstd{nb}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{textmodel_nb}\hlstd{(both_dfms}\hlopt{$}\hlstd{train,} \hlkwc{y}\hlstd{=quanteda}\hlopt{::}\hlkwd{docvars}\hlstd{(both_dfms}\hlopt{$}\hlstd{train,} \hlstr{"EV_article"}\hlstd{),} \hlkwc{prior}\hlstd{=}\hlstr{"uniform"}\hlstd{)}
\hlstd{prob_nb}\hlkwb{<-}\hlkwd{predict}\hlstd{(nb,} \hlkwc{newdata} \hlstd{= both_dfms}\hlopt{$}\hlstd{test,} \hlkwc{type}\hlstd{=}\hlstr{"probability"}\hlstd{)}
\hlstd{pred_nb}\hlkwb{<-}\hlkwd{data.frame}\hlstd{(}\hlkwd{predict}\hlstd{(nb,} \hlkwc{newdata} \hlstd{= both_dfms}\hlopt{$}\hlstd{test,} \hlkwc{type}\hlstd{=}\hlstr{"class"}\hlstd{))}
\hlstd{res}\hlkwb{<-}\hlkwd{data.frame}\hlstd{(}\hlkwc{predict_nb}\hlstd{=pred_nb, prob_nb)}
\hlkwd{names}\hlstd{(res)}\hlkwb{<-}\hlkwd{c}\hlstd{(}\hlstr{"predict_nb"}\hlstd{,} \hlstr{"prob_notev"}\hlstd{,} \hlstr{"prob_ev"}\hlstd{)}
\hlkwd{assess_classification}\hlstd{(}\hlkwd{organise_results}\hlstd{(both_dfms}\hlopt{$}\hlstd{test, res))} \hlopt{%>%}
  \hlkwd{filter}\hlstd{(rowname} \hlopt{%in%} \hlkwd{c}\hlstd{(}\hlstr{"Precision"}\hlstd{,} \hlstr{"Recall"}\hlstd{,} \hlstr{"F1"}\hlstd{))} \hlopt{%>%}
  \hlkwd{kable}\hlstd{()}
\end{alltt}
\end{kframe}
\begin{tabular}{l|r|l}
\hline
rowname & value & model\\
\hline
Precision & 0.7842105 & naive bayes\\
\hline
Recall & 0.8370787 & naive bayes\\
\hline
F1 & 0.8097826 & naive bayes\\
\hline
\end{tabular}

\begin{kframe}\begin{alltt}
\hlstd{description_corpus}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{corpus}\hlstd{(classdocs[,}\hlkwd{c}\hlstd{(}\hlstr{"fakeid"}\hlstd{,} \hlstr{"description"}\hlstd{,} \hlstr{"EV_article"}\hlstd{)],} \hlkwc{text_field} \hlstd{=} \hlstr{"description"}\hlstd{)}
\hlstd{description_dfm} \hlkwb{<-} \hlkwd{preprocess_corpus}\hlstd{(description_corpus,} \hlkwc{stem}\hlstd{=}\hlnum{FALSE}\hlstd{,} \hlkwc{min_termfreq}\hlstd{=}\hlnum{5}\hlstd{,} \hlkwc{min_docfreq} \hlstd{=} \hlnum{2}\hlstd{)}

\hlstd{both_dfms}\hlkwb{<-}\hlkwd{split_dfm}\hlstd{(description_dfm,} \hlkwc{n_train} \hlstd{=} \hlnum{1000}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

Note: this way of creating the dfm does make dfms terms equal because one overall dfm is created and then it is subset. You can see below that the number of features in both dfms is the same:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{print}\hlstd{(both_dfms}\hlopt{$}\hlstd{train)}
\end{alltt}
\begin{verbatim}
## Document-feature matrix of: 1,000 documents, 1,647 features (99.2% sparse).
\end{verbatim}
\begin{alltt}
\hlkwd{print}\hlstd{(both_dfms}\hlopt{$}\hlstd{test)}
\end{alltt}
\begin{verbatim}
## Document-feature matrix of: 894 documents, 1,647 features (99.2% sparse).
\end{verbatim}
\end{kframe}
\end{knitrout}

\subsection{More realistic use case}
Exclude non-election cases which would not be returned by our keywords:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{description_corpus}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{corpus}\hlstd{(classdocs[classdocs}\hlopt{$}\hlstd{election_article}\hlopt{==}\hlnum{1}\hlstd{,}\hlkwd{c}\hlstd{(}\hlstr{"fakeid"}\hlstd{,} \hlstr{"description"}\hlstd{,} \hlstr{"EV_article"}\hlstd{)],} \hlkwc{text_field} \hlstd{=} \hlstr{"description"}\hlstd{)}
\hlstd{description_dfm} \hlkwb{<-} \hlkwd{preprocess_corpus}\hlstd{(description_corpus,} \hlkwc{stem}\hlstd{=}\hlnum{FALSE}\hlstd{,} \hlkwc{min_termfreq}\hlstd{=}\hlnum{5}\hlstd{,} \hlkwc{min_docfreq} \hlstd{=} \hlnum{2}\hlstd{)}

\hlstd{both_dfms}\hlkwb{<-}\hlkwd{split_dfm}\hlstd{(description_dfm,} \hlkwc{n_train} \hlstd{=} \hlnum{1000}\hlstd{)}

\hlstd{nb}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{textmodel_nb}\hlstd{(both_dfms}\hlopt{$}\hlstd{train,} \hlkwc{y}\hlstd{=quanteda}\hlopt{::}\hlkwd{docvars}\hlstd{(both_dfms}\hlopt{$}\hlstd{train,} \hlstr{"EV_article"}\hlstd{),} \hlkwc{prior}\hlstd{=}\hlstr{"uniform"}\hlstd{)}
\hlstd{prob_nb}\hlkwb{<-}\hlkwd{predict}\hlstd{(nb,} \hlkwc{newdata} \hlstd{= both_dfms}\hlopt{$}\hlstd{test,} \hlkwc{type}\hlstd{=}\hlstr{"probability"}\hlstd{)}
\hlstd{pred_nb}\hlkwb{<-}\hlkwd{data.frame}\hlstd{(}\hlkwd{predict}\hlstd{(nb,} \hlkwc{newdata} \hlstd{= both_dfms}\hlopt{$}\hlstd{test,} \hlkwc{type}\hlstd{=}\hlstr{"class"}\hlstd{))}
\hlstd{res}\hlkwb{<-}\hlkwd{data.frame}\hlstd{(}\hlkwc{predict_nb}\hlstd{=pred_nb, prob_nb)}
\hlkwd{names}\hlstd{(res)}\hlkwb{<-}\hlkwd{c}\hlstd{(}\hlstr{"predict_nb"}\hlstd{,} \hlstr{"prob_notev"}\hlstd{,} \hlstr{"prob_ev"}\hlstd{)}
\hlkwd{assess_classification}\hlstd{(}\hlkwd{organise_results}\hlstd{(both_dfms}\hlopt{$}\hlstd{test, res))} \hlopt{%>%}
  \hlkwd{filter}\hlstd{(rowname} \hlopt{%in%} \hlkwd{c}\hlstd{(}\hlstr{"Precision"}\hlstd{,} \hlstr{"Recall"}\hlstd{,} \hlstr{"F1"}\hlstd{))} \hlopt{%>%}
  \hlkwd{kable}\hlstd{()}
\end{alltt}
\end{kframe}
\begin{tabular}{l|r|l}
\hline
rowname & value & model\\
\hline
Precision & 0.9019608 & naive bayes\\
\hline
Recall & 0.7076923 & naive bayes\\
\hline
F1 & 0.7931034 & naive bayes\\
\hline
\end{tabular}

\begin{kframe}\begin{alltt}
\hlstd{description_corpus}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{corpus}\hlstd{(classdocs[,}\hlkwd{c}\hlstd{(}\hlstr{"fakeid"}\hlstd{,} \hlstr{"description"}\hlstd{,} \hlstr{"EV_article"}\hlstd{)],} \hlkwc{text_field} \hlstd{=} \hlstr{"description"}\hlstd{)}
\hlstd{description_dfm} \hlkwb{<-} \hlkwd{preprocess_corpus}\hlstd{(description_corpus,} \hlkwc{stem}\hlstd{=}\hlnum{FALSE}\hlstd{,} \hlkwc{min_termfreq}\hlstd{=}\hlnum{5}\hlstd{,} \hlkwc{min_docfreq} \hlstd{=} \hlnum{2}\hlstd{)}

\hlstd{both_dfms}\hlkwb{<-}\hlkwd{split_dfm}\hlstd{(description_dfm,} \hlkwc{n_train} \hlstd{=} \hlnum{1000}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\section{Refinement Using OCR}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{ocr_corpus}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{corpus}\hlstd{(classdocs[,}\hlkwd{c}\hlstd{(}\hlstr{"fakeid"}\hlstd{,} \hlstr{"ocr"}\hlstd{,} \hlstr{"EV_article"}\hlstd{)],} \hlkwc{text_field} \hlstd{=} \hlstr{"ocr"}\hlstd{)}
\hlstd{ocr_dfm} \hlkwb{<-} \hlkwd{preprocess_corpus}\hlstd{(ocr_corpus,} \hlkwc{stem}\hlstd{=}\hlnum{FALSE}\hlstd{,} \hlkwc{min_termfreq}\hlstd{=}\hlnum{5}\hlstd{,} \hlkwc{min_docfreq} \hlstd{=} \hlnum{2}\hlstd{)}

\hlstd{both_dfms}\hlkwb{<-}\hlkwd{split_dfm}\hlstd{(ocr_dfm,} \hlkwc{n_train} \hlstd{=} \hlnum{1000}\hlstd{)}
\hlstd{nb}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{textmodel_nb}\hlstd{(both_dfms}\hlopt{$}\hlstd{train,} \hlkwc{y}\hlstd{=quanteda}\hlopt{::}\hlkwd{docvars}\hlstd{(both_dfms}\hlopt{$}\hlstd{train,} \hlstr{"EV_article"}\hlstd{),} \hlkwc{prior}\hlstd{=}\hlstr{"uniform"}\hlstd{)}
\hlstd{prob_nb}\hlkwb{<-}\hlkwd{predict}\hlstd{(nb,} \hlkwc{newdata} \hlstd{= both_dfms}\hlopt{$}\hlstd{test,} \hlkwc{type}\hlstd{=}\hlstr{"probability"}\hlstd{)}
\hlstd{pred_nb}\hlkwb{<-}\hlkwd{data.frame}\hlstd{(}\hlkwd{predict}\hlstd{(nb,} \hlkwc{newdata} \hlstd{= both_dfms}\hlopt{$}\hlstd{test,} \hlkwc{type}\hlstd{=}\hlstr{"class"}\hlstd{))}
\hlstd{res}\hlkwb{<-}\hlkwd{data.frame}\hlstd{(}\hlkwc{predict_nb}\hlstd{=pred_nb, prob_nb)}
\hlkwd{names}\hlstd{(res)}\hlkwb{<-}\hlkwd{c}\hlstd{(}\hlstr{"predict_nb"}\hlstd{,} \hlstr{"prob_notev"}\hlstd{,} \hlstr{"prob_ev"}\hlstd{)}
\hlkwd{assess_classification}\hlstd{(}\hlkwd{organise_results}\hlstd{(both_dfms}\hlopt{$}\hlstd{test, res))} \hlopt{%>%}
  \hlkwd{filter}\hlstd{(rowname} \hlopt{%in%} \hlkwd{c}\hlstd{(}\hlstr{"Precision"}\hlstd{,} \hlstr{"Recall"}\hlstd{,} \hlstr{"F1"}\hlstd{))} \hlopt{%>%}
  \hlkwd{kable}\hlstd{()}
\end{alltt}
\end{kframe}
\begin{tabular}{l|r|l}
\hline
rowname & value & model\\
\hline
Precision & 0.6456212 & naive bayes\\
\hline
Recall & 0.9031339 & naive bayes\\
\hline
F1 & 0.7529691 & naive bayes\\
\hline
\end{tabular}

\end{knitrout}

\section{King Algorithm: implementation}

\subsection{Incrementally Define $R$ and $S$}
$R$ is our reference set. [King suggestions: could define $R$ based on one simple keyword search].
\subsubsection{Intermediate step}
Take keywords in $R$, $K_R$, ranked by simple statistic such as document frequency or frequency-inverse document frequency. User examines elements of $K_R$ apart from those used to define the set and chooses some keywords to define $Q_S$, which in turn generates a definition for $S$ so that we can run the rest of the algorithm.
The user can continue to add keywords from $K_R$ into the final desired query $Q_RT$.
This step also mitigates the issue of how to define a search set in large data sets that do not fit into memory all at once or may not even be able to be retrieved all at onece. is the BNA.

\subsection{Partition $S$ into $T$ and $S \backslash T$}
To partition $S$ into $T$ and $S \backslash T$, we first define a `training' set by sampling from $S$ and $R$. Since $R$ is typically much smaller than $S$ our test set for our classifiers is all all of $S$, we often use the entire $R$ set and a sample of $S$ as our training set.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{R} \hlkwb{<-} \hlstd{classdocs[classdocs}\hlopt{$}\hlstd{EV_article}\hlopt{==}\hlnum{1}\hlstd{,]}
\hlstd{R}\hlopt{$}\hlstd{R} \hlkwb{<-} \hlnum{1}
\hlstd{R}\hlopt{$}\hlstd{in_sample}\hlkwb{<-}\hlnum{1}
\hlstd{S} \hlkwb{<-} \hlstd{unclassdocs}
\hlstd{S}\hlopt{$}\hlstd{R} \hlkwb{<-} \hlnum{0}

\hlstd{n_sample_S} \hlkwb{<-} \hlnum{1000}
\hlstd{S}\hlopt{$}\hlstd{in_sample}\hlkwb{<-}\hlnum{0}
\hlstd{S}\hlopt{$}\hlstd{in_sample[}\hlkwd{sample}\hlstd{(}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(S), n_sample_S)]}\hlkwb{<-}\hlnum{1}

\hlstd{R_S} \hlkwb{<-} \hlstd{dplyr}\hlopt{::}\hlkwd{bind_rows}\hlstd{(R, S)}
\hlstd{R_S}\hlopt{$}\hlstd{fakeid}\hlkwb{<-}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(R_S)}

\hlstd{R_S_corpus}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{corpus}\hlstd{(R_S[,}\hlkwd{c}\hlstd{(}\hlstr{"fakeid"}\hlstd{,} \hlstr{"ocr"}\hlstd{,} \hlstr{"in_sample"}\hlstd{,} \hlstr{"R"}\hlstd{)],} \hlkwc{text_field} \hlstd{=} \hlstr{"ocr"}\hlstd{)}
\hlstd{R_S_dfm} \hlkwb{<-} \hlkwd{preprocess_corpus}\hlstd{(R_S_corpus,} \hlkwc{stem}\hlstd{=}\hlnum{FALSE}\hlstd{,} \hlkwc{min_termfreq}\hlstd{=}\hlnum{20}\hlstd{,} \hlkwc{min_docfreq} \hlstd{=} \hlnum{20}\hlstd{)}

\hlstd{king_train_dfm}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{dfm_subset}\hlstd{(R_S_dfm, quanteda}\hlopt{::}\hlkwd{docvars}\hlstd{(R_S_dfm,} \hlstr{"in_sample"}\hlstd{)}\hlopt{==}\hlnum{1}\hlstd{)}
\hlstd{king_nb} \hlkwb{<-} \hlstd{quanteda}\hlopt{::}\hlkwd{textmodel_nb}\hlstd{(king_train_dfm,} \hlkwc{y}\hlstd{=quanteda}\hlopt{::}\hlkwd{docvars}\hlstd{(king_train_dfm,} \hlstr{"R"}\hlstd{),} \hlkwc{prior}\hlstd{=}\hlstr{"uniform"}\hlstd{)}
\hlstd{keywords}\hlkwb{<-}\hlkwd{nb_keywords}\hlstd{(king_train_dfm,} \hlstr{"R"}\hlstd{)}
\hlstd{knitr}\hlopt{::}\hlkwd{kable}\hlstd{(}\hlkwd{head}\hlstd{(keywords,} \hlnum{20}\hlstd{))}
\end{alltt}
\end{kframe}
\begin{tabular}{l|l|r|r|r}
\hline
  & rowname & 0 & 1 & id\\
\hline
3 & generated & 0.0221304 & 0.9778696 & 3\\
\hline
14 & legible & 0.0227857 & 0.9772143 & 14\\
\hline
8 & technology & 0.0250073 & 0.9749927 & 8\\
\hline
13 & deciphering & 0.0250073 & 0.9749927 & 13\\
\hline
2919 & chartist & 0.0267458 & 0.9732542 & 2919\\
\hline
2670 & detective & 0.0337936 & 0.9662064 & 2670\\
\hline
2271 & roughs & 0.0350241 & 0.9649759 & 2271\\
\hline
9415 & telegraphed & 0.0353459 & 0.9646541 & 9415\\
\hline
12921 & gutted & 0.0370477 & 0.9629523 & 12921\\
\hline
2501 & p.c & 0.0449904 & 0.9550096 & 2501\\
\hline
1079 & cobden & 0.0458847 & 0.9541153 & 1079\\
\hline
3421 & smashing & 0.0495804 & 0.9504196 & 3421\\
\hline
10653 & worker & 0.0558817 & 0.9441183 & 10653\\
\hline
9331 & schoolroom & 0.0579874 & 0.9420126 & 9331\\
\hline
13218 & gladstone's & 0.0579874 & 0.9420126 & 13218\\
\hline
3510 & ratepayers & 0.0602579 & 0.9397421 & 3510\\
\hline
8408 & telegram & 0.0632289 & 0.9367711 & 8408\\
\hline
1167 & disraeli & 0.0653778 & 0.9346222 & 1167\\
\hline
5 & optical & 0.0714484 & 0.9285516 & 5\\
\hline
12923 & pontypool & 0.0787619 & 0.9212381 & 12923\\
\hline
\end{tabular}

\end{knitrout}

After fitting the classifiers, we use the estimated parameters to generate predicted probabilities of $R$ membership for all documents in $S$. Of cource, all the search set documents in fact fall within $S$ but our interest in in learning from the \textit{mistakes} these classifiers make.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{S_dfm} \hlkwb{<-} \hlstd{quanteda}\hlopt{::}\hlkwd{dfm_subset}\hlstd{(R_S_dfm, quanteda}\hlopt{::}\hlkwd{docvars}\hlstd{(R_S_dfm,} \hlstr{"R"}\hlstd{)}\hlopt{==}\hlnum{0}\hlstd{)}
\hlstd{quanteda}\hlopt{::}\hlkwd{docvars}\hlstd{(S_dfm,} \hlstr{"T"}\hlstd{)}\hlkwb{<-}\hlkwd{predict}\hlstd{(king_nb,} \hlkwc{newdata} \hlstd{= S_dfm,} \hlkwc{type}\hlstd{=}\hlstr{"class"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\subsection{Discovering Keywords to Classify Documents}
After partitioning $S$ into estimated target $T$ and non-target set $S \backslash T$, we must find and rank keywords which best discriminate $T$ and $S \backslash T$.
King does this in three stages:
\subsubsection{identify keywords in S}

\subsubsection{sort them into those that predict each of the two steps}

My suggestion here is simply to use binary\_dfm

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{S_dfm_binary}\hlkwb{<-}\hlstd{quanteda}\hlopt{::}\hlkwd{dfm_weight}\hlstd{(S_dfm,} \hlkwc{scheme}\hlstd{=}\hlstr{"boolean"}\hlstd{)}
\hlstd{king_stage2} \hlkwb{<-} \hlstd{quanteda}\hlopt{::}\hlkwd{textmodel_nb}\hlstd{(S_dfm,} \hlkwc{y}\hlstd{=quanteda}\hlopt{::}\hlkwd{docvars}\hlstd{(S_dfm_binary,} \hlstr{"T"}\hlstd{),} \hlkwc{distribution}\hlstd{=}\hlstr{"Bernoulli"}\hlstd{)}
\hlstd{king_stage2} \hlkwb{<-} \hlkwd{nb_keywords}\hlstd{(S_dfm,} \hlstr{"T"}\hlstd{)}
\hlstd{knitr}\hlopt{::}\hlkwd{kable}\hlstd{(}\hlkwd{head}\hlstd{(king_stage2,} \hlnum{30}\hlstd{))}
\end{alltt}
\end{kframe}
\begin{tabular}{l|l|r|r|r}
\hline
  & rowname & 0 & 1 & id\\
\hline
8722 & unionism & 0.0113223 & 0.9886777 & 8722\\
\hline
8429 & horden & 0.0141129 & 0.9858871 & 8429\\
\hline
1208 & protectionist & 0.0173133 & 0.9826867 & 1208\\
\hline
9203 & suffragists & 0.0203970 & 0.9796030 & 9203\\
\hline
9590 & nationalist & 0.0223910 & 0.9776090 & 9590\\
\hline
8485 & motor & 0.0237863 & 0.9762137 & 8485\\
\hline
10142 & socialism & 0.0290080 & 0.9709920 & 10142\\
\hline
8420 & nationalists & 0.0296335 & 0.9703665 & 8420\\
\hline
9858 & socialist & 0.0296335 & 0.9703665 & 9858\\
\hline
2335 & dump & 0.0302866 & 0.9697134 & 2335\\
\hline
9120 & bicycle & 0.0316832 & 0.9683168 & 9120\\
\hline
9011 & truncheons & 0.0399786 & 0.9600214 & 9011\\
\hline
3074 & asquith & 0.0427787 & 0.9572213 & 3074\\
\hline
8737 & industries & 0.0438013 & 0.9561987 & 8737\\
\hline
9286 & rowdyism & 0.0438013 & 0.9561987 & 9286\\
\hline
9331 & schoolroom & 0.0438013 & 0.9561987 & 9331\\
\hline
9415 & telegraphed & 0.0438013 & 0.9561987 & 9415\\
\hline
8994 & batons & 0.0455433 & 0.9544567 & 8994\\
\hline
8421 & socialists & 0.0460006 & 0.9539994 & 8421\\
\hline
2417 & workers & 0.0471852 & 0.9528148 & 2417\\
\hline
8747 & candidature & 0.0491752 & 0.9508248 & 8747\\
\hline
2631 & unionist & 0.0495718 & 0.9504282 & 2631\\
\hline
9138 & railings & 0.0502017 & 0.9497983 & 9138\\
\hline
9205 & barricaded & 0.0502017 & 0.9497983 & 9205\\
\hline
1993 & witness's & 0.0541586 & 0.9458414 & 1993\\
\hline
2919 & chartist & 0.0541586 & 0.9458414 & 2919\\
\hline
8609 & drafted & 0.0541586 & 0.9458414 & 8609\\
\hline
9043 & booing & 0.0541586 & 0.9458414 & 9043\\
\hline
8398 & hartlepool & 0.0575614 & 0.9424386 & 8398\\
\hline
8430 & sacked & 0.0598163 & 0.9401837 & 8430\\
\hline
\end{tabular}

\begin{kframe}\begin{alltt}
\hlstd{knitr}\hlopt{::}\hlkwd{kable}\hlstd{(}\hlkwd{tail}\hlstd{(king_stage2,} \hlnum{30}\hlstd{))}
\end{alltt}
\end{kframe}
\begin{tabular}{l|l|r|r|r}
\hline
  & rowname & 0 & 1 & id\\
\hline
16564 & despised & 0.9361369 & 0.0638631 & 16564\\
\hline
97 & fruits & 0.9370575 & 0.0629425 & 97\\
\hline
16555 & breaching & 0.9370575 & 0.0629425 & 16555\\
\hline
5001 & candidly & 0.9379520 & 0.0620480 & 5001\\
\hline
16375 & fortress & 0.9379520 & 0.0620480 & 16375\\
\hline
4632 & citadel & 0.9381717 & 0.0618283 & 4632\\
\hline
16237 & scrip & 0.9388215 & 0.0611785 & 16237\\
\hline
7969 & lunette & 0.9392471 & 0.0607529 & 7969\\
\hline
4023 & aspire & 0.9420685 & 0.0579315 & 4023\\
\hline
7970 & laurent & 0.9428271 & 0.0571729 & 7970\\
\hline
13935 & shells & 0.9449882 & 0.0550118 & 13935\\
\hline
15303 & carolina & 0.9449882 & 0.0550118 & 15303\\
\hline
6762 & combat & 0.9463404 & 0.0536596 & 6762\\
\hline
7376 & batteries & 0.9504040 & 0.0495960 & 7376\\
\hline
4633 & belgians & 0.9537326 & 0.0462674 & 4633\\
\hline
14316 & belgian & 0.9556141 & 0.0443859 & 14316\\
\hline
7397 & repose & 0.9560607 & 0.0439393 & 7397\\
\hline
4619 & antwerp & 0.9562807 & 0.0437193 & 4619\\
\hline
11134 & senate & 0.9564985 & 0.0435015 & 11134\\
\hline
12926 & prussia & 0.9564985 & 0.0435015 & 12926\\
\hline
2897 & gerard & 0.9566425 & 0.0433575 & 2897\\
\hline
15893 & frontiers & 0.9577611 & 0.0422389 & 15893\\
\hline
4618 & capitulation & 0.9581658 & 0.0418342 & 4618\\
\hline
16545 & lillo & 0.9611442 & 0.0388558 & 16545\\
\hline
6394 & belgium & 0.9623721 & 0.0376279 & 6394\\
\hline
4623 & chasse & 0.9626404 & 0.0373596 & 4623\\
\hline
7977 & bombs & 0.9672625 & 0.0327375 & 7977\\
\hline
7981 & scheldt & 0.9679827 & 0.0320173 & 7981\\
\hline
7973 & besiegers & 0.9727743 & 0.0272257 & 7973\\
\hline
4625 & forts & 0.9769438 & 0.0230562 & 4625\\
\hline
\end{tabular}

\end{knitrout}

\subsection{rank them by degree of discriminatory power}
We already seem to have done this in the step above
\subsection{Human Input and Human-Computer Iteration}
Now present

\end{document}