### 2018-08-02 other_R_code/article_selection_pilot_writeup.Rnw
add article selection pilot code

\documentclass{article}
\usepackage{amsmath}

\begin{document}

<<setuppath, echo=FALSE, results="hide">>=
dropbox.path<-switch (Sys.info()["nodename"],
                      "DM-GIA-051"="D:/Dropbox",
                      "DM-GIA-055"="D:/Dropbox",
                      "GIDON-ASUS-15"="C:/Users/Gidon/Dropbox",
                      "GID-HOME-LENOVO"="C:/Users/Gidon/Dropbox")

project.path<-paste0(dropbox.path, "/ESRC Grant EV 19th Century")

knitr::read_chunk(paste0(project.path, "/Code/durhamevp/other_R_code/article_selection_pilot.R"))
@

<<pilot.packages, warning=FALSE, message=FALSE>>=

@

<<download.R.set, cache=TRUE>>=

@

<<initial.1832.results>>=

@

\end{document}