### 2018-08-09 other_R_code/article_selection_pilot_evaluate_keyword_methods.Rnw
implementation pilot functions

\documentclass{article}
\usepackage[margin=1in]{geometry}
\title{Election Violence Newspaper Article Selection Process Pilot: Evaluate Keyword Selection Method}
\begin{document}
\maketitle
\section{Summary}
After the code to setup and download data (Sections \ref{sect:packages}-\ref{sect:download1832}) the document below includes:
First, an implementation of code to select of articles for coding, where initial evaluation of articles is based on keywords from the boolean searches which return those articles (Sections \ref{sect:whichdownloadocr}-\ref{sect:useonocr}). This classification is done using actual searches conducted (effectively 4 keywords) with a Naive Bayes classifier.

Second, the document then provides an evaluation of the keyword selection step based on user classified articles from the platform with Naive Bayes and other classifiers and with 4 and 10 keyword sets (Sections \ref{sect:eval4nb}-\ref{sect:eval10other}).

Finally, to provide an alternative way of evaluating the process, the document includes 10 randomly selected cases which are rejected for coding in the implementation code (Section \ref{sect:rejectnbprocess}).

\section{Packages used}
\label{sect:packages}
<<setuppath, echo=FALSE, results="hide">>=
dropbox.path<-switch (Sys.info()["nodename"],
                      "DM-GIA-051"="D:/Dropbox",
                      "DM-GIA-055"="D:/Dropbox",
                      "GIDON-ASUS-15"="C:/Users/Gidon/Dropbox",
                      "GID-HOME-LENOVO"="C:/Users/Gidon/Dropbox")

project.path<-paste0(dropbox.path, "/ESRC Grant EV 19th Century")

knitr::read_chunk(paste0(project.path, "/Code/durhamevp/other_R_code/article_selection_pilot.R"))
options(tidy=TRUE, width=50)
@

<<resetdb, echo=FALSE, warning=FALSE, message=FALSE, results="hide">>=
# Is this necessary because of cached results below? Otherwise sometime get error messages like Corrupt MySQL handle
durhamevp::killDbConnections()

@

<<selection.pilot.custom.functions, echo=FALSE>>=

@

<<pilot.packages, warning=FALSE, message=FALSE>>=

@

\section{Download Classified Cases}
There is an initial set of cases which represent the concept (in this case `election violence'). We also have a set of cases which do not represent the concept (Note: King et al do not have the non-election violence cases).

I also add indicators that this material is classified (\texttt{classified==1}) and that it is not unclassified (\texttt{unclass==0}) which will be useful when added to subsequent material.
<<download.R.set, cache=TRUE>>=

@

\section{Assess Naive Bayes selection performance on 4 keywords (binary dfm)}
\label{sect:eval4nb}
Here there are four keywords (election, riot, disturbance, incident) which are either present (1) or not-present (0). Use these patterns to identify election violence articles with a Naive Bayes classifer.
<<subset.on.keywords.eval.four.words>>=

@

\section{How do other classifiers perform on 4 keywords (binary dfm)}
Is performance improved by using other classifiers than Naive Bayes? Here there are four keywords (election, riot, disturbance, incident) which are either present (1) or not-present (0). Use these patterns to identify election violence articles.
<<subset.on.keywords.eval.other.methods.four.words>>=

@

\section{Assess Naive Bayes selection performance on 10 keywords (binary dfm)}
Here there are ten keywords ((election, riot, disturbance, incident, mob, stone, window, candidate, party, hustings, magistrate) which are either present (1) or not-present (0). Use these patterns to identify election violence articles with a Naive Bayes classifer.
<<subset.on.keywords.eval.ten.words>>=

@

\section{How do other classifiers perform on 10 keywords (binary dfm)}
\label{sect:eval10other}
The same but with ten keywords (election, riot, disturbance, incident, mob, stone, window, candidate, party, hustings, magistrate) which are either present (1) or not-present (0). Use these patterns to identify election violence articles using a range of other classifiers.

<<subset.on.keywords.eval.other.methods.ten.words>>=

@

\end{document}