### 2018-08-14 other_R_code/article_selection_pilot_description_examples.Rnw
update xgboost/nb test

\documentclass{article}
\usepackage[margin=1in]{geometry}
\title{Low Hanging Fruit: Descriptions of 1841 from the results returned by the new keyword searches}
\begin{document}
\maketitle

\label{sect:packages}
<<setuppath, echo=FALSE, results="hide", echo=FALSE>>=
dropbox.path<-switch (Sys.info()["nodename"],
                      "DM-GIA-051"="D:/Dropbox",
                      "DM-GIA-055"="D:/Dropbox",
                      "GIDON-ASUS-15"="C:/Users/Gidon/Dropbox",
                      "GID-HOME-LENOVO"="C:/Users/Gidon/Dropbox")

project.path<-paste0(dropbox.path, "/ESRC Grant EV 19th Century")

knitr::read_chunk(paste0(project.path, "/Code/durhamevp/other_R_code/article_selection_pilot.R"))
options(tidy=TRUE, width=50)
@

<<resetdb, echo=FALSE, results='hide', message=FALSE>>=
# Is this necessary because of cached results below? Otherwise sometime get error messages like Corrupt MySQL handle
durhamevp::killDbConnections()

@

<<selection.pilot.custom.functions, echo=FALSE>>=

@

<<pilot.packages, warning=FALSE, message=FALSE, echo=FALSE>>=

@

<<resetdbagain, echo=FALSE, results='hide', message=FALSE>>=
# Is this necessary because of cached results below? Otherwise sometime get error messages like Corrupt MySQL handle
durhamevp::killDbConnections()

@

<<download.R.set, cache=TRUE, echo=FALSE>>=

@

<<download.candocs, cache=TRUE, echo=FALSE>>=

@

<<subset.on.description.1841.data, cache=TRUE, echo=FALSE, message=FALSE, results='hide'>>=

@

Below are two randomly selected sets of articles from the two classificaiton methods. Which is the better set of articles?
\section{First Set}
<<printdoc1, results='asis', echo=FALSE>>=
doc1<-dplyr::sample_n(select_descript_xgb, 30)

documents_to_latex(doc1, include_ocr=FALSE)
@

\section{Second Set}
<<printdoc2, results='asis', echo=FALSE>>=
doc2<-dplyr::sample_n(select_descript_nb, 30)
documents_to_latex(doc2, include_ocr=FALSE)

@

\end{document}