### 2018-09-11 other_R_code/user_activity_gdh.Rmd
Gary's sandbox version of user activity document, creating user status report - for learning use only

---
title: "User Status Report"
author: "Gidon Cohen"
date: "14 June 2018"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)

library(durhamevp)
library(tidyverse)
library(knitr)
killDbConnections <- function () {

  all_cons <- DBI::dbListConnections(RMySQL::MySQL())

  print(all_cons)

  for(con in all_cons) DBI::dbDisconnect(con)

  print(paste(length(all_cons), " connections killed."))

}
killDbConnections()
con<-evdb_connect(password_method = "keyring")
activity<-DBI::dbGetQuery(con, 'select *
                from portal_useractivitylogentry')
doc_allocations<-DBI::dbGetQuery(con, 'select *
                from portal_userdocumentallocation')
documents<-DBI::dbGetQuery(con, 'select *
                from portal_document')
event_report<-DBI::dbGetQuery(con, 'select *
                from portal_eventreport')

```

## Time Taken on Documents

```{r timetaken, echo=FALSE}

inputPanel(
  selectInput("allocation_type", label = "Mode:",
              choices = c("training", "testing", "coding")),
    selectInput("facet_by", label = "Facet by:",
              choices = c("user_id", "document_id"))
)

renderPlot({
  doc_allocations$facet_col <- paste(input$facet_by, doc_allocations[,input$facet_by])

  activity %>%
    left_join(doc_allocations, by=c("allocation_id"="id", "user_id")) %>%
    filter(user_id>7) %>%
    filter(allocation_type==input$allocation_type) %>%
    filter(status=="COMPLETED") %>%
    group_by(user_id, allocation_id, facet_col) %>%
    summarize(start=min(date), end=max(date), time_taken=difftime(end, start)) %>%
    ggplot(aes(time_taken))+
    facet_wrap(~facet_col)+
    geom_histogram()

})
```

## Are Documents in Coding Mode Relevant to Election Violence?

```{r, echo=FALSE, results='asis'}

aa <- doc_allocations %>%
  filter(user_id>7) %>%
  filter(status=="COMPLETED") %>%
  filter(allocation_type=="coding") %>%
  group_by(violent_nature, violent_focus, electoral_nature, geo_relevant, time_relevant) %>%
  tally()

kable(aa)
```

## Is Time Taken in Coding Mode Related to Difficulty and/or Word Count?

```{r echo=FALSE}

time_summary<- activity %>%
  left_join(doc_allocations, by=c("allocation_id"="id", "user_id")) %>%
  left_join(documents, by=c("document_id"="id")) %>%
  filter(user_id>7) %>%
  filter(allocation_type=="coding") %>%
  filter(word_count>0) %>%
  filter(status=="COMPLETED")

time_summary %>%
  group_by(user_id, allocation_id, word_count, violent_nature) %>%
  summarize(start=min(date), end=max(date), time_taken=difftime(end, start), difficulty_ranking=mean(difficulty_ranking)) %>%
  ggplot(aes(time_taken))+
  facet_wrap(~violent_nature)+
  geom_histogram()

time_summary %>%
  group_by(user_id, allocation_id, word_count) %>%
  summarize(start=min(date), end=max(date), time_taken=difftime(end, start), difficulty_ranking=mean(difficulty_ranking)) %>%
  ggplot(aes(difficulty_ranking, time_taken))+
  #ggplot(aes(difficulty_ranking, time_taken))+
  geom_point(position = position_jitter())+
  stat_smooth(method="lm", se=FALSE, alpha=.2, colour="black", aes(group=user_id))+
  stat_smooth(method="lm", se=TRUE)

time_summary %>%
  filter(word_count>0) %>%
  group_by(user_id, allocation_id, word_count) %>%
  summarize(start=min(date), end=max(date), time_taken=difftime(end, start), difficulty_ranking=mean(difficulty_ranking)) %>%
  ggplot(aes(log(word_count), time_taken))+
  geom_point()+
  stat_smooth(method="lm", se=TRUE
  )

```