### 2018-09-21 R/Reporting/coder_matching_report.Rmd
Commit to change branch

---
title: "User Matched Scoring Report"
author: "Gidon Cohen"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
library(magrittr)
library(tidyverse)
dropbox.path<-switch (Sys.info()["nodename"],
                      "DM-GIA-051"="D:/Dropbox",
                      "DM-GIA-055"="D:/Dropbox",
                      "GIDON-ASUS-15"="C:/Users/Gidon/Dropbox",
                      "GID-HOME-LENOVO"="C:/Users/Gidon/Dropbox")

project.path<-paste0(dropbox.path, "/ESRC Grant EV 19th Century")
match_data_path<-paste0(project.path, "/Code/database_interactivity/reporting/coder_scoring/data")

knitr::read_chunk(paste0(project.path, "/Code/database_interactivity/reporting/coder_scoring/analyse_matches.R"))
```

```{r readdata, echo=FALSE, warning=FALSE}

```

## Overall Summary of Matching by Variable

```{r overallagreement, echo=FALSE, warning=FALSE}

```

## Individual Users Contribution to Coding Agreement (Document Allocation Variables)

```{r linearuserdoc, echo=FALSE, warning=FALSE}

```

## Individual Users Contribution to Coding Agreement (All Variables)

```{r linearall, echo=FALSE, warning=FALSE}

```

## Summary of agreement by user

```{r agreementsummarybyuser, results='asis', warning=FALSE, echo=FALSE}

```